\sys: Light-touch Data-Parallel Shell Processing

### 2020-10-09 paper/draft.tex
Remove some space above equations
Add updated program numbers in eval, slightly improve abstract, address Caleb's comment for PoSh

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
This paper presents \sys, a system aimed at parallelizing POSIX shell scripts through a combination of program transformations and runtime primitives.
% a source-to-source shell compiler and associated runtime library
Given a script, \sys converts it to a dataflow graph, performs a series of analyses and transformations that aim at exposing parallelism while preserving its sequential semantics, and then converts the dataflow graph back into a script that uses POSIX constructs to explicitly guide parallelism.
A lightweight annotation language allow command developers to express key parallelizability properties about their commands.
An accompanying parallelizability study of POSIX and GNU commands---two large and commonly used groups---informs \sys's annotation language and guides the development of a data-parallel standard library.
% \kk{eurosys: Can we extend this to say that PaSh automatically uses constructs from this library? Otherwise it sounds like the user has to use it to make their scripts parallel.}
Finally, \sys's \unix-aware runtime primitives address several practical issues related to performance and correctness.
\sys's extensive evaluation, combining several classic and modern \unix scripts, shows significant \todo{speedups} stemming from the combination of its program transformations and runtime primitives.
\end{abstract}

@@

We evaluate \sys using
  (i) a series of benchmarks, ranging from classic \unix one-liners to modern data-processing scripts;
  (ii) two large and complex use cases for temperature analysis and web indexing;
%% , and container orchestration;
and 
  (iii) a series of micro-benchmarks inspired by (and comparing with) possible alternatives.
The vast majority (47/52) of scripts---all unmodified---see significant speedups.
Indicatively, applied to 10 representative \unix one-liners,
%  with \ttt{--width=64}, 
 \sys achieves 9.35$\times$ due to program transformations alone and 13.24$\times$ when coupled with its runtime primitives.
% \kk{eurosys: Readers don't know what --width is here so it might make sense to not mention it.}
% \sys performs better than \ttt{sort --parallel}, a hand-tuned parallel implementation, and GNU parallel, which returns incorrect results if used without care.
When applied to a large program used for temperature analysis, 
  \sys successfully accelerates it by 2.52$\times$---
  parallelizing both the computation as well as the preprocessing fragment (downloading, extracting, and cleaning the data) which traditionally falls outside of the focus of conventional parallelization systems even though it takes $75\%$ of the whole program's execution time.
  % extracting 12.31$\times$ from the portion of the program that traditionally falls outside the focus of conventional parallelization systems.
  % \kk{\sys parallelizes the computation by 12.31 and not the preprocessing. It parallelizes the preprocessing by 2.}

@@

\heading{Stateless and Parallelizable Pure Commands}
Stateless commands such as \ttt{tr}
operate independently on individual elements of the stream
(characters, lines, or files) without maintaining any state~\sx{cmd}. In this
work, we consider lines as the data quantum, thus we consider stateless only commands that are
stateless with respect to lines (or any finer granularity such as
characters). Formally, a streaming command $f$ is stateless if it commutes with
the operation of concatenation, \ie it is a semigroup homomorphism:
%% with respect to concatenation \km{It commutes with the operation of concatenation. This is the definition of (semigroup) homomorphism. I am not sure that linearity is an appropriate term.}
%% , \ie satisfying the following equation:

\[
\forall x, x', s, f(x \cdot x', s) = f(x, s) \cdot f(x', s)
\]

%% This is a nice characterization, which I think is true in most
%% settings. There is a caveat though: You can have stateless programs
%% that emit output when they start (before they consume the stream)
%% and/or when they consume the EOF symbol. In this case I am not sure
%% the equation above holds.

\noindent
This means that applying the command $f$ to a concatenation of two
inputs $x, x'$ produces the same output as applying $f$ to each input
$x, x'$ separately, and concatenating the outputs.

Similarly, some pure commands (such as \ttt{sort} and \ttt{wc}) can be
parallelized using divide-and-conquer techniques. More formally, these
pure commands $f$ can be implemented as a combination of a function map
$m$ and an associative aggregate $agg$, satisfying the following equation:

\[
\forall x, x', s, f(x \cdot x', s) = agg(m(x, s) \cdot m(x', s), s)
\]

\noindent
This means that we get the same output by applying $f$ to a
concatenation of two inputs $x, x'$ as by applying the aggregation function
$agg$ to the concatenation of the outputs produced by applying map $m$
to each of $x$ and $x'$.
%% \km{We had also discussed here the case of
%%   operators that can be described as associative and commutative
%%   functions on the data (e.g., sum). In this case the parallelization
%%   has the flavor of data parallelism: one can split the input stream
%%   in an arbitrary way (e.g., randomly or using some other
%%   load-balancing scheme) and merge the outputs. For the more broad
%%   class of associative operations (e.g., sort), the formulation of
%%   this paragraph makes perfect sense. I just want to mention that this
%%   feels a lot more like \textbf{batching} than map-reduce-like data
%%   parallelization.} \kk{Thanks, I think you are right. Batching is a
%%   better name for it. In the end we didn't handle
%%   commutative-associative function (that are more like map-reduce) and
%%   so we changes the name to map-fold to not imply commutativity.}

@@

\heading{Highlights}
This paragraph highlights results for \ttt{width=16}, but \sys's evaluation reports on varying \ttt{width}s (2--64).
Overall, \sys accelerates 47/52 scripts---all unmodified---by 1.92--17.42$\times$; 
  for the rest, the parallel performance is comparable to the sequential (0.89, 0.91, 0.94, 099, 1.01$\times$).
  % \kk{eurosys: I have an issue that I am not sure how to address. I would like to have the 61.1x here (which is due to 64 width)  but I don't see a good way of including it without sounding suspicious.}
For a subset of our benchmarks~\sx{ours}, we decouple the speedup due to \sys's DFG transformations alone from the one achieved when including its runtime primitives:
  the former averages $5.9\times$, jumping to $8.6\times$ due to the latter.
A different subset ($30$ benchmarks) sees average speedups of $6.02\times$~\sx{unix50}.
% , depending on the classed of commands involved
Two complex scripts with 15~\sx{macro1} and 34~\sx{macro2} commands written in multiple languages see speedups of $2.44\times$ and $12.7\times$.
\sys significantly outperforms \ttt{sort --parallel}, a hand-tuned parallel implementation, and performs better than GNU \ttt{parallel}, which returns incorrect results if used without care~\sx{micro}.
% Averages: [1.955223776839714, 3.478591807121036, 5.721328556478535, 8.604380509595906, 10.666921423438112, 13.246864076708423]
% No Eager Averages: [1.6327834370824543, 2.5481043878162444, 3.8623841529643173, 5.931851620007484, 7.466566647963421, 9.35752032439642]
% From the rest, only two see a slowdown, only because \sys's setup ($\sim1s$) is higher than their runtime ($\sim0.4s$).
% \sys's optimized primitives offer benefits even against manually parallelized scripts. % (such as GNU sort, and GNU parallel).
% \sys's compilation pass is negligible (avg. $<1s$). %  for non trivial program execution.
% In two large cases, tasks such as data download, extraction, and preprocessing---often implemented using shell scripts and outside the focus of specialized parallelization frameworks---take comparable time to the execution of the main computation, and can be significantly accelerated by \sys.
% In terms of manual effort,

Using \sys's standard library of annotations for POSIX and GNU commands~\sx{parallelizability}, the vast majority of programs ($>50$, with $>200$ commands) require no effort to parallelize other than invoking \sys;
  only 6 ($<3\%$) commands, outside this library, needed a single-record annotation~\sx{macro2}.

@@

\heading{Results}
The complete pipeline executes in 44m2s for five years (82GB) of data.
\sys with \ttt{width=}16 leads to 2.52$\times$ speedup, with different phases seeing different benefits:
  2.04$\times$ (vs. 33m58s) for all the pre-processing and 
  12.31$\times$ speedup (vs. 10m4s) for computing the max.

@@

Concurrent to \sys, POSH~\cite{raghavan2020posh} is a framework for running shell scripts over remote storage by offloading the I/O-intensive portions to proxy servers closer to the data.
POSH focuses on co-locating distributed operation, and introduces a novel scheduling algorithm. 
\sys instead focuses on the parallelization of CPU-intensive scripts, by transforming them to DFGs, applying transformations, and then transforming them back to parallel schell scripts augmented with \sys's runtime primitives to be executed on a standard shell.
%  non-distributed parallelism, is location-agnostic, and focuses on correctness through its dataflow transformations.

@@

This paper presented \sys, a system allowing shell users to parallelize shell programs mostly automatically.
% a source-to-source shell compiler and associated runtime library
\sys can be thought as a parallelizing source-to-source compiler coupled with a runtime component.
The compiler operates by converting dataflow regions into dataflow graphs and then applying iterative transformations based on the characteristics of commands in these regions.
Characteristics are known by the  use of classification and annotations (for certain classes), which \sys offers for commands in POSIX and Coreutils used frequently.
\sys's \unix-aware runtime primitives address several practical issues related to performance and correctness.
Applying \sys to classic and modern \unix scripts shows significant speedups---in one set of benchmarks, the compiler's average speedup is $9.3\times$, jumping to $13.2\times$ due to runtime optimizations.
\sys is open source and available for download at \ttt{blinded}.

>>>>

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
This paper presents \sys, a system for parallelizing POSIX shell scripts through a combination of program transformations and runtime primitives.
% a source-to-source shell compiler and associated runtime library
Given a script, \sys converts it to a dataflow graph, performs a series of transformations that expose parallelism while preserving its sequential semantics, and then converts the dataflow graph back into a script that uses POSIX constructs to explicitly guide parallelism.
A lightweight annotation language allows command developers to express key parallelizability properties about their commands.
An accompanying parallelizability study of POSIX and GNU commands---two large and commonly used groups---informs said annotation language and guides the development of an optimized aggregator library that \sys uses.
% \kk{eurosys: Can we extend this to say that PaSh automatically uses constructs from this library? Otherwise it sounds like the user has to use it to make their scripts parallel.}
Finally, \sys's \unix-aware runtime primitives address several practical issues related to performance and correctness.
\sys's extensive evaluation, combining several classic and modern \unix scripts, shows significant speedups (total average over 44 scripts is $6.7\times$) stemming from the combination of its program transformations and runtime primitives.
% The vast majority (39/44) of scripts---all unmodified---see significant speedups (average over all is $6.7\times$, up to $61\times$).
\end{abstract}

@@

We evaluate \sys using
  (i) a series of benchmarks, ranging from classic \unix one-liners to modern data-processing scripts;
  (ii) two large and complex use cases for temperature analysis and web indexing;
%% , and container orchestration;
and 
  (iii) a series of micro-benchmarks inspired by (and comparing with) possible alternatives.
The vast majority (39/44) of scripts---all unmodified---see significant speedups (average over all is $6.7\times$, up to $61\times$).
Indicatively, applied to 10 representative \unix one-liners,
%  with \ttt{--width=64}, 
 \sys achieves 9.35$\times$ due to program transformations alone and 13.24$\times$ when coupled with its runtime primitives.
% \kk{eurosys: Readers don't know what --width is here so it might make sense to not mention it.}
% \sys performs better than \ttt{sort --parallel}, a hand-tuned parallel implementation, and GNU parallel, which returns incorrect results if used without care.
When applied to a large program used for temperature analysis, 
  \sys successfully accelerates it by 2.52$\times$---
  parallelizing both the computation as well as the preprocessing fragment (downloading, extracting, and cleaning the data) which traditionally falls outside of the focus of conventional parallelization systems even though it takes $75\%$ of the whole program's execution time.
  % extracting 12.31$\times$ from the portion of the program that traditionally falls outside the focus of conventional parallelization systems.
  % \kk{\sys parallelizes the computation by 12.31 and not the preprocessing. It parallelizes the preprocessing by 2.}

@@

\heading{Stateless and Parallelizable Pure Commands}
Stateless commands such as \ttt{tr}
operate independently on individual elements of the stream
(characters, lines, or files) without maintaining any state~\sx{cmd}. In this
work, we consider lines as the data quantum, thus we consider stateless only commands that are
stateless with respect to lines (or any finer granularity such as
characters). Formally, a streaming command $f$ is stateless if it commutes with
the operation of concatenation, \ie it is a semigroup homomorphism:
%% with respect to concatenation \km{It commutes with the operation of concatenation. This is the definition of (semigroup) homomorphism. I am not sure that linearity is an appropriate term.}
%% , \ie satisfying the following equation:
%
\[
\forall x, x', s, f(x \cdot x', s) = f(x, s) \cdot f(x', s)
\]
%
%% This is a nice characterization, which I think is true in most
%% settings. There is a caveat though: You can have stateless programs
%% that emit output when they start (before they consume the stream)
%% and/or when they consume the EOF symbol. In this case I am not sure
%% the equation above holds.
%
\noindent
This means that applying the command $f$ to a concatenation of two
inputs $x, x'$ produces the same output as applying $f$ to each input
$x, x'$ separately, and concatenating the outputs.

Similarly, some pure commands (such as \ttt{sort} and \ttt{wc}) can be
parallelized using divide-and-conquer techniques. More formally, these
pure commands $f$ can be implemented as a combination of a function map
$m$ and an associative aggregate $agg$, satisfying the following equation:
%
\[
\forall x, x', s, f(x \cdot x', s) = agg(m(x, s) \cdot m(x', s), s)
\]
%
\noindent
This means that we get the same output by applying $f$ to a
concatenation of two inputs $x, x'$ as by applying the aggregation function
$agg$ to the concatenation of the outputs produced by applying map $m$
to each of $x$ and $x'$.
%% \km{We had also discussed here the case of
%%   operators that can be described as associative and commutative
%%   functions on the data (e.g., sum). In this case the parallelization
%%   has the flavor of data parallelism: one can split the input stream
%%   in an arbitrary way (e.g., randomly or using some other
%%   load-balancing scheme) and merge the outputs. For the more broad
%%   class of associative operations (e.g., sort), the formulation of
%%   this paragraph makes perfect sense. I just want to mention that this
%%   feels a lot more like \textbf{batching} than map-reduce-like data
%%   parallelization.} \kk{Thanks, I think you are right. Batching is a
%%   better name for it. In the end we didn't handle
%%   commutative-associative function (that are more like map-reduce) and
%%   so we changes the name to map-fold to not imply commutativity.}

@@

\heading{Highlights}
This paragraph highlights results for \ttt{width=16}, but \sys's evaluation reports on varying \ttt{width}s (2--64).
% The vast majority (39/44) of scripts---all unmodified---see significant speedups (average over all is $6.7\times$, up to $61\times$).
Overall, \sys accelerates 39/44 scripts---all unmodified---by 1.92--17.42$\times$; 
  for the rest, the parallel performance is comparable to the sequential (0.89, 0.91, 0.94, 099, 1.01$\times$).
  % \kk{eurosys: I have an issue that I am not sure how to address. I would like to have the 61.1x here (which is due to 64 width)  but I don't see a good way of including it without sounding suspicious.}
For a subset of our benchmarks~\sx{ours}, we decouple the speedup due to \sys's DFG transformations alone from the one achieved when including its runtime primitives:
  the former averages $5.9\times$, jumping to $8.6\times$ due to the latter.
A different subset ($30$ benchmarks) sees average speedups of $6.02\times$~\sx{unix50}.
% , depending on the classed of commands involved
Two complex scripts with 15~\sx{macro1} and 34~\sx{macro2} commands written in multiple languages see speedups of $2.44\times$ and $12.7\times$.
\sys significantly outperforms \ttt{sort --parallel}, a hand-tuned parallel implementation, and performs better than GNU \ttt{parallel}, which returns incorrect results if used without care~\sx{micro}.
% Averages: [1.955223776839714, 3.478591807121036, 5.721328556478535, 8.604380509595906, 10.666921423438112, 13.246864076708423]
% No Eager Averages: [1.6327834370824543, 2.5481043878162444, 3.8623841529643173, 5.931851620007484, 7.466566647963421, 9.35752032439642]
% From the rest, only two see a slowdown, only because \sys's setup ($\sim1s$) is higher than their runtime ($\sim0.4s$).
% \sys's optimized primitives offer benefits even against manually parallelized scripts. % (such as GNU sort, and GNU parallel).
% \sys's compilation pass is negligible (avg. $<1s$). %  for non trivial program execution.
% In two large cases, tasks such as data download, extraction, and preprocessing---often implemented using shell scripts and outside the focus of specialized parallelization frameworks---take comparable time to the execution of the main computation, and can be significantly accelerated by \sys.
% In terms of manual effort,

Using \sys's standard library of annotations for POSIX and GNU commands~\sx{parallelizability}, the vast majority of programs ($>40$, with $>200$ commands) require no effort to parallelize other than invoking \sys;
  only 6 ($<3\%$) commands, outside this library, needed a single-record annotation~\sx{macro2}.

@@

\heading{Results}
The complete pipeline executes in 44m2s for five years (82GB) of data.
\sys with \ttt{width=}16 leads to 2.52$\times$ speedup, with different phases seeing different benefits:
  2.04$\times$ (vs. 33m58s) for all the pre-processing ($75\%$ of the total running time) and 
  12.31$\times$ speedup (vs. 10m4s) for computing the max.

@@

Developed independently and at the same time with \sys, POSH~\cite{raghavan2020posh} is a framework for running shell scripts over remote storage by offloading the I/O-intensive portions to proxy servers closer to the data.
POSH focuses on co-locating distributed operation, and introduces a novel scheduling algorithm. 
\sys instead focuses on the parallelization of CPU-intensive scripts, by transforming them to DFGs, applying transformations, and then transforming them back to parallel schell scripts augmented with \sys's runtime primitives to be executed on a standard shell.
%  non-distributed parallelism, is location-agnostic, and focuses on correctness through its dataflow transformations.

@@

This paper presented \sys, a system allowing shell users to parallelize shell programs mostly automatically.
% a source-to-source shell compiler and associated runtime library
\sys can be thought as a parallelizing source-to-source compiler coupled with a runtime component.
The compiler operates by converting dataflow regions into dataflow graphs and then applying iterative transformations based on the characteristics of commands in these regions.
Characteristics are known by the  use of classification and annotations (for certain classes), which \sys offers for commands in POSIX and Coreutils used frequently.
\sys's \unix-aware runtime primitives address several practical issues related to performance and correctness.
Our evaluation of \sys on a wide variety of \emph{unmodified} scripts shows that the vast majority (39/44) see significant speedups (average over all is $6.7\times$, up to $61\times$).
% Applying \sys to classic and modern \unix scripts shows significant speedups---in one set of benchmarks, the compiler's average speedup is $9.3\times$, jumping to $13.2\times$ due to runtime optimizations.
\sys is open source and available for download online~[Blind].
